---
title: "experiment5"
author: "Aske & Rebecca"
date: "6 October 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

EXPERIMENT 5 - Discriminability

Load packages
```{r}
# packages
pacman::p_load(tidyverse,lmerTest,ggbeeswarm, brms, rethinking)

```


Collect data from several csv files and write a new csv file
```{r}

# WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp5_data")


# import data from several csv files
fileList5 <- list.files(full.names = T, pattern =".csv")

d5 <- map_dfr(fileList5, read_csv) %>% 
  mutate(
    Participant = as.factor(Participant),
    Age = as.factor(Age),
    Gender = as.factor(Gender),
    Trial = as.factor(Trial),
    Stimulus1 = as.factor(Stimulus1),
    Stimulus2 = as.factor(Stimulus2),
    Target = as.factor(Target),
    Response = as.factor(Response)
  )


# Create csv in shared folder
write.csv(d5,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp5_collected_data.csv", row.names = FALSE)


```


# Import collected data
```{r}
# WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske")

# Import collected data
d5_old <- read_csv("exp5_collected_data.csv")

## For ID 56-60

# set WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp5_data/1-6")

# import data from several csv files
fileList5.2 <- list.files(full.names = T, pattern =".csv")
list2 <- lapply(fileList5.2, read_csv)
d5.2 <- do.call(rbind, list2)


```

Gather old df and 1-6
```{r}
# change participant number before merging
d5.2$Participant <- as.numeric(d5.2$Participant)
d5.2$Participant <- d5.2$Participant + 55

# rbind them
d5 <- rbind(d5_old,d5.2)

# Create csv in shared folder
write.csv(d5,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp5_collected_data_preprocessed.csv", row.names = FALSE)

```

Flip the the 'correct column' in the cases, where it is clearly opposite
```{r}

setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske")
d5 <- read_csv("exp5_collected_data_preprocessed.csv")

# summarise showing that there are IDs with flipped 0's and 1's in Correct column
correctness <- d5 %>% group_by(Participant) %>% 
  summarise(
    correctness = mean(Correct)
  )
correctness


# Make list of problematic IDs
fails <- correctness %>% filter(correctness < 0.2)

# Create empty df with correct rows
d5_clean <- d5[0,]

# Loop correcting the 0's and 1's.
for (id in 1:length(unique(d5$Participant))){
  
  # Make subset for each participant
  ID <- dplyr::filter(d5, Participant == id)
  
  
  if (id %in% c(fails$Participant) == TRUE){
    
    ID$Correct[ID$Correct == 0] <- 0.9
    ID$Correct[ID$Correct == 1] <- 0.1
    
    ID$Correct[ID$Correct == 0.1] <- 0
    ID$Correct[ID$Correct == 0.9] <- 1
    
  }
  
  # Combine with premade empty dataframe
       if (nrow(d5_clean) == 0) {
         d5_clean <- ID
         } else {
         d5_clean <- rbind(d5_clean, ID)}
  
}



# Save the old data in a placeholder 
d5_old <- d5

# overwrite d5 with the new data
d5 <- d5_clean


```

Create chainGroup and StimPair columns
```{r}

# Change to factors
d5$Correct <- as.factor(d5$Correct)
# No reaction times under 0.2
d5 <- subset(d5,Reaction_time > 0.2)



### CREATE chainGroup column - Loop through and add chain group ID
d5$Participant <- as.numeric(d5$Participant)

d5$chainGroup <- NA
for (i in (1:nrow(d5))){
  
  if (d5$Participant[i] <= 5) {
    d5$chainGroup[i] <- 1
    
  } else if (d5$Participant[i] > 5 & d5$Participant[i] < 11 | d5$Participant[i] == 61) {
    d5$chainGroup[i] <- 2
    
  } else if (d5$Participant[i] > 10 & d5$Participant[i] < 16) {
    d5$chainGroup[i] <- 3
    
  } else if (d5$Participant[i] > 15 & d5$Participant[i] < 21) {
    d5$chainGroup[i] <- 4
    
  } else if (d5$Participant[i] > 20 & d5$Participant[i] < 26) {
    d5$chainGroup[i] <- 5
    
  } else if (d5$Participant[i] > 25 & d5$Participant[i] < 31) {
    d5$chainGroup[i] <- 6
    
  } else if (d5$Participant[i] > 30 & d5$Participant[i] < 36) {
    d5$chainGroup[i] <- 7
    
  } else if (d5$Participant[i] > 35 & d5$Participant[i] < 41) {
    d5$chainGroup[i] <- 8
    
  } else if (d5$Participant[i] > 40 & d5$Participant[i] < 46) {
    d5$chainGroup[i] <- 9
    
  } else if (d5$Participant[i] > 45 & d5$Participant[i] < 51) {
    d5$chainGroup[i] <- 10
    
  } else if (d5$Participant[i] > 50 & d5$Participant[i] < 56) {
    d5$chainGroup[i] <- 11
    
  } else { # In case we add 56-61 (which had the pictures from group 1)
    d5$chainGroup[i] <- 1
  }
}

# as factor
d5$Participant <- as.factor(d5$Participant)
d5$chainGroup <- as.factor(d5$chainGroup)

# make orientation column with values constrained between 0 and 1
d5$OriL = d5$Ori/max(d5$Ori)



### Create StimPairID column 

d5_new <- d5[0,] # make empty dataframe
count = 1 # create count

# the LOOP - add StimPairID column
for (group in 1:length(unique(d5$chainGroup))){
  
  
  ## for each chainGroup loop through and create StimPair unique ID
  sub <- subset(d5, chainGroup == group)



  # Create list of target pictures in the chain group (all 24 pics are represented in at least one participant)
  stimuli <- unique(sub$Target)


  x <- NULL
  x <- data.frame(matrix(unlist(combn(stimuli, 2, simplify=F)), nrow=276, byrow=T)) # 276 = (24*24-24)/24
  x$StimPairID <- seq(nrow(x))

  # Loop adding stimPair value to the current subset
  sub$StimPairID <- NA
  
  for (i in seq(nrow(x))){
    sub$StimPairID[
      (sub$Stimulus1 %in% x$X1[i] | sub$Stimulus1 %in% x$X2[i]) & 
      (sub$Stimulus2 %in% x$X1[i] | sub$Stimulus2 %in% x$X2[i])] = x$StimPairID[i]
  }

  
  # Add value so they all become unique across stim groups
  sub$StimPairID <- sub$StimPairID + (1000 * count)
  count <- count + 1
  
  
  # Combine with premade empty dataframe
  if (nrow(d5_new) == 0) {
    d5_new <- sub
    } else {
        d5_new <- rbind(d5_new, sub)
  }
  
}

# Check 
check <- d5_new %>% group_by(Participant) %>% summarise(
  count = n(),
  unikke = length(unique(StimPairID))
)


# Overwrite d5
d5 <- d5_new

```

Create within chain dummy column and generationNUmber column
```{r}

# within chain
d5$Stim1Chain <- substr(d5$Stimulus1, start = 5, stop = 7)
d5$Stim2Chain <- substr(d5$Stimulus2, start = 5, stop = 7)

d5$withinChain <- ifelse(d5$Stim1Chain == d5$Stim2Chain, 1, 0)


# generationNumber column
# Change values of generation in new colum to be -1, 0 or 1.
d5$generationNumber <- as.numeric(d5$Generation_target)

d5$generationNumber <- ifelse(d5$generationNumber == 1, -1, d5$generationNumber)
d5$generationNumber <- ifelse(d5$generationNumber == 4, 0, d5$generationNumber)
d5$generationNumber <- ifelse(d5$generationNumber == 8, 1, d5$generationNumber)


# Save the final dataframe
write.csv(d5,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp5_collected_data_preprocessed_FINAL.csv", row.names = FALSE)

```


# Modelling

Remove bad pics
```{r}
# Import the completely pre-processed data
d5_allPics <- read_csv("exp5_collected_data_preprocessed_FINAL.csv")

# subset bad pics
problemPics <- c("i3_c104_g8_id16", "i3_c104_g4_id12", "i3_c104_g1_id9",
                 "i4_c102_g4_id15", "i4_c102_g8_id315", "i4_c102_g1_",
                 "i4_c108_g4_id48", "i4_c108_g8_id52", "i4_c108_g1_id45",
                 "i3_c103_g8_id8", "i3_c103_g4_id4", "i3_c103_g1_id1")

d5 <- d5_allPics %>% 
  subset(!Stimulus1 %in% problemPics) %>% 
  subset(!Stimulus2 %in% problemPics) %>% 
  subset(!Target %in% problemPics)


d5 <- d5 %>% mutate(
  Participant = as.factor(Participant),
  Age = as.factor(Age),
  Gender = as.factor(Gender),
  Stimulus1 = as.factor(Stimulus1),
  Stimulus2 = as.factor(Stimulus2),
  Target = as.factor(Target),
  chainGroup = as.factor(chainGroup),
  Age = as.factor(Age),
  StimPairID = as.factor(StimPairID)
)

d5_between <- subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0)
d5_within <- subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1)


```


## Between chain
```{r}

# set values for running bayesian models
CHAINS = 4
CORES = 4
ITER = 4000
WARMUP = 1000


# Formula5.1 - linear
formula5.1 <- bf(Reaction_time ~ generationNumber + 
                   (1 + generationNumber|Participant) + 
                   (1|Target) + 
                   (1|StimPairID))

# Formula5.2 - quadratic
formula5.2 <- bf(Reaction_time ~ generationNumber + I(generationNumber^2) + 
                   (1 + generationNumber + I(generationNumber^2)|Participant) + 
                   (1|Target) + 
                   (1|StimPairID))


# Set priors
get_prior(formula5.1, d5, family = lognormal())

prior5 <- c(
  prior(normal(0.5, 0.2), class = Intercept),
  prior(normal(0, 0.5), class = b),
  prior(normal(0, 0.2), class = sd),
  prior(normal(0, 0.2), class = sigma),
  prior(lkj(5), class = cor))

### MODEL A1 - Linear

# prior
m5.1a_prior <- brm(
  formula5.1,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.1a_prior, nsamples=100)

# model
m5.1a <- brm(
  formula5.1,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# post check
pp_check(m5.1a, nsamples=100)




### MODEL A2 - Quadratic

# prior
m5.2a_prior <- brm(
  formula5.2,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.2a_prior, nsamples=100)

# post
m5.2a <- brm(
  formula5.2,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# post check
pp_check(m5.2a, nsamples=100)

```


### Results
```{r}

# Summaries
summary(m5.1a)
summary(m5.2a)


# model LOOIC comparison
m5.1a <- add_criterion(m5.1a, criterion = c("bayes_R2", "loo"))
m5.2a <- add_criterion(m5.2a, criterion = c("bayes_R2", "loo"))

exp5_compare1 <- loo_compare(m5.1a, m5.2a)
exp5_weights1 <- loo_model_weights(m5.1a, m5.2a)

exp5_compare1
exp5_weights1

# hypothesis testing
hypothesis(m5.1a, "generationNumber < 0") 
plot(hypothesis(m5.1a, "generationNumber < 0"))

hypothesis(m5.2a, "generationNumber < 0") 
plot(hypothesis(m5.2a, "generationNumber < 0"))


# remember trace plots etc.
plot(m5.1a)
plot(m5.2a)


```

the simple linear model was found most credible --> we run analysis controlling for orientation on it.


## Within chain
```{r}

### MODEL B1 - Linear

#prior
m5.1b_prior <- brm(
  formula5.1,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.1b_prior, nsamples=100)


m5.1b <- brm(
  formula5.1,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# prior predictive check
pp_check(m5.1b, nsamples=100)



### MODEL B2 - Quadratic

# prior
m5.2b_prior <- brm(
  formula5.2,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.2b_prior, nsamples=100)

# model
m5.2b <- brm(
  formula5.2,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# prior predictive check
pp_check(m5.2b, nsamples=100)


```


### Results
```{r}

# Summaries
summary(m5.1b)
summary(m5.2b)


# model LOOIC comparison
m5.1b <- add_criterion(m5.1b, criterion = c("bayes_R2", "loo"))
m5.2b <- add_criterion(m5.2b, criterion = c("bayes_R2", "loo"))

exp5_compare2 <- loo_compare(m5.1b, m5.2b)
exp5_weights2 <- loo_model_weights(m5.1b, m5.2b)

exp5_weights2


# hypothesis testing
hypothesis(m5.1b, "generationNumber < 0") 
plot(hypothesis(m5.1b, "generationNumber < 0"))

hypothesis(m5.2b, "generationNumber < 0") 
plot(hypothesis(m5.2b, "generationNumber < 0"))


# remember trace plots etc.
plot(m5.1b)
plot(m5.2b)


```


## Control for mental rotation by letting generation interact with OriL
```{r}

# define formula
formula5.3 <- bf(Reaction_time ~ generationNumber * OriL + 
                    (1 + generationNumber * OriL | Participant) + 
                    (1 + OriL|Target) + 
                    (1 + OriL|StimPairID))


# Set priors
get_prior(formula5.3, d5, family = lognormal())


# A - between sites

#prior
m5.3a_prior <- brm(
  formula5.3,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.3a_prior, nsamples=100)


m5.3a <- brm(
  formula5.3,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 0),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# prior predictive check
pp_check(m5.3a, nsamples=100)



# B - within sites

# prior
m5.3b_prior <- brm(
  formula5.3,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = "only",
  chains = 1,
  cores = 1,
  iter = 200)

# prior predictive check
pp_check(m5.3b_prior, nsamples=100)

# model
m5.3b <- brm(
  formula5.3,
  subset(d5, Same_generation == 1 & Correct==1 & withinChain == 1),
  family = lognormal(),
  prior = prior5,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER)

# prior predictive check
pp_check(m5.3b, nsamples=100)



```


### Results
```{r}

# Summaries
summary(m5.3a)
summary(m5.3b)


# model LOOIC comparison
m5.3a <- add_criterion(m5.3a, criterion = c("bayes_R2", "loo"))
m5.3b <- add_criterion(m5.3b, criterion = c("bayes_R2", "loo"))

exp5_compare3 <- loo_compare(m5.1a, m5.2a, m5.3a)
exp5_weights3 <- loo_model_weights(m5.1a, m5.2a, m5.3a)

exp5_compare4 <- loo_compare(m5.1b, m5.2b, m5.3b)
exp5_weights4 <- loo_model_weights(m5.1b, m5.2b, m5.3b)

exp5_weights3
exp5_compare3
# hypothesis testing


# remember trace plots etc.
plot(m5.3a)
plot(m5.3b)


# Hypothesis testinbg
hypothesis(m5.3a, "generationNumber < 0")
hypothesis(m5.3a, "OriL > 0")

hypothesis(m5.3b, "generationNumber < 0")
hypothesis(m5.3b, "OriL > 0")


# Re calculate estimates in seconds to enhance interpretatibility

# across chains
exp(-0.22) # 1st generation
exp(-0.22-0.08) # 4th generation
exp(-0.22-0.08-0.08) # 8th generation

# Within chains
exp(-0.20) # 1st generation
exp(-0.20-0.1) # 4th generation
exp(-0.20-0.1-0.1) # 8th generation

```








# Plots
```{r}

# BETWEEN CHAIN (A)
plotSum5a <- d5 %>% filter(withinChain == 0) %>% group_by(Participant, generationNumber) %>% summarise(
  RTmean = mean(Reaction_time)
)

plotSum5a$Participant <- as.factor(plotSum5a$Participant)
plotSum5a$generationNumber <- as.numeric(plotSum5a$generationNumber)
library(font)
p_load(extrafont, font)
loadfonts(device = "win")
windowsFonts(Times = windowsFont("TT Times New Roman"))

theme_set(theme_bw(base_size=12, base_family='Times New Roman'))

Exp5a_MainPlot <- ggplot(plotSum5a, aes(generationNumber, RTmean)) + 
  geom_line(aes(group = Participant, color = Participant), alpha = 0.6) +
  geom_point(aes(group = Participant, color = Participant), alpha = 0.6) +
  geom_smooth(method=lm, color = "red") +
  scale_color_discrete(guide=FALSE) +
  scale_x_continuous(breaks = c(-1,0,1), labels=c("1st","4th","8th")) +

  # THEMES
  theme_grey() +
  theme(
    text = element_text(family = "Times New Roman"),
    legend.position="top",
    plot.subtitle=element_text(face="italic",size=14,colour="grey40"),
    plot.title=element_text(size=21,face="bold")) +
  labs(title="\nExperiment 5a",
       subtitle="Discriminability across Chains",
       x=expression("Generation"),
       y=expression("Time to Discriminate (seconds)")) +
  NULL
  
Exp5a_MainPlot

### WITHIN CHAIN (B)
plotSum5b <- d5 %>% filter(withinChain == 1) %>% group_by(Participant, generationNumber) %>% summarise(
  RTmean = mean(Reaction_time)
)

plotSum5b$Participant <- as.factor(plotSum5b$Participant)
plotSum5b$generationNumber <- as.numeric(plotSum5b$generationNumber)

Exp5b_MainPlot <- ggplot(plotSum5b, aes(generationNumber, RTmean)) + 
  geom_line(aes(group = Participant, color = Participant), alpha = 0.6) +
  geom_point(aes(group = Participant, color = Participant), alpha = 0.6) +
  geom_smooth(method=lm, color = "red") +
  scale_color_discrete(guide=FALSE) +
  scale_x_continuous(breaks = c(-1,0,1), labels=c("1st","4th","8th")) + 
  
  # THEMES
  theme_grey() +
  theme(
    text = element_text(family = "Times New Roman"),
    legend.position="top",
    plot.subtitle=element_text(face="italic",size=14,colour="grey40"),
    plot.title=element_text(size=21,face="bold")) +
  labs(title="\nExperiment 5b",
       subtitle="Discriminability within Chains",
       x=expression("Generation"),
       y=expression("Time to Discriminate (seconds)")) +
  NULL

  
Exp5b_MainPlot


```




































