---
title: "aske_rebecca_experiment3"
author: "Aske & Rebecca"
date: "7 October 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# Load packages
library(pacman)
p_load(tidyverse, brms, ggbeeswarm, loo)


# set WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp3_data")

# import data from several csv files
fileList3 <- list.files(full.names = T, pattern =".csv")
list <- lapply(fileList3, read_csv)
d3 <- do.call(rbind, list)

# Create csv in shared folder
write.csv(d3,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp3_collected_data.csv", row.names = FALSE)




## For ID 56-60

# set WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp3_data/1-6")

# import data from several csv files
fileList3.2 <- list.files(full.names = T, pattern =".csv")
list2 <- lapply(fileList3.2, read_csv)
d3.2 <- do.call(rbind, list2)


```


Check data
```{r}
# set wd
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske")

d3_old <- read_csv("exp3_collected_data.csv")

# Check datapoints per participants
check <- d3_old %>% group_by(drawing, generation) %>% summarize(n())
names(check)[3] <- "count"
check %>% filter(count != 60)
# Correct number of rows for each generation

check <- d3_old %>% group_by(ID, drawing, generation) %>% summarize(n())
names(check)[4] <- "count"
check %>% filter(count != 6)
# Correct number of rows for each drawing/ID/generation
check

```


Merge with similarity scores
```{r}

# load data
similarityScores <- read_csv("SimilarityData.csv")

# remove 0's in ID column
similarityScores$ID = str_remove(similarityScores$ID, "^0+")

# merge
d3 <- merge(d3_old, similarityScores, by=c("ID","drawing")) 

# check
check <- d3 %>% group_by(ID) %>% summarise(
  pictures = length(unique(drawing))
)


# remove unnecesary columns
d3$X1.x <- NULL
d3$X1.y <- NULL


```

Make new df for the last 6 participants (incl. similarity scores)
```{r}
# load data
similarityScores_1to5 <- read_csv("SimilarityData_1to5.csv")

# merge
d3.2_merged <- merge(d3.2, similarityScores_1to5, by=c("ID","drawing")) 

# check
d3.2_merged %>% group_by(ID) %>% summarise(
  pictures = length(unique(drawing))
)


# remove unnecesary columns
d3.2_merged$X1.x <- NULL
d3.2_merged$X1.y <- NULL

```

Gather the 2 df
```{r}

# change participant number before merging
d3.2_merged$ID <- d3.2_merged$ID + 55

# rbind them
d3 <- rbind(d3,d3.2_merged)

# Create csv in shared folder
write.csv(d3,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp3_collected_data_preprocessed.csv", row.names = FALSE)

```


# ANALYSIS

Prepare parameters
```{r}

d3 <- read_csv("exp3_collected_data_preprocessed.csv")

# standardize generation parameter to be -1, 0, and 1.
d3$generation <- factor(d3$generation, levels = c(1, 4, 8))
d3$generationL = as.numeric(d3$generation)
d3$generationL <- d3$generationL - 2

# Transform MSE so that it is on a 0-1 scale
d3$MSE_s = d3$MSE - min(d3$MSE)
d3$MSE_s = d3$MSE_s / max(d3$MSE_s)

# fairly normally distributed
hist(d3$MSE_s)

```


# Modelling

## Predicting error MSE from generation

linear and quadratic models (prior and posterior)
```{r}

# define chains, iter, and controls
CHAINS = 4
CORES = 2
ITER = 4e3

CONTROLS = list(
  max_treedepth = 20,
  adapt_delta=0.99)

# make subset
sub <- c(6, 7, 8, 9, 10)


# Construct models
formula3.1 <- bf(MSE_s ~ 1 + generationL + (1 + generationL | ID) + (1 | drawing))
formula3.2 <- bf(MSE_s ~ 1 + generationL + I(generationL)^2 + (1 + generationL + I(generationL)^2 | ID) + (1 | drawing))

# get priors to be set
get_prior(formula3.1, d3, family = "gaussian")

# set priors
prior <- c(
  prior(normal(0,.3), class = Intercept),
  prior(normal(0,.1), class = b),
  prior(normal(0,.05), class = sd),
  prior(lkj(5), class = cor)
)




### MODEL 1

# Run model based on priors alone
m3.1_prior <- brm(
   formula3.1,
   data = subset(d3, ID %in% sub),
   family = gaussian,
   prior = prior,
   sample_prior = "only",
   iter = ITER,
   warmup = 1000,
   chains = CHAINS,
   cores = 1,
   control = CONTROLS
)

# prior predictive check
pp_check(m3.1_prior, nsamples=100)

# Run model
m3.1 <- brm(
   formula3.1,
   data = d3, # subset(d3, ID %in% sub)
   family = gaussian,
   prior = prior,
   sample_prior = TRUE,
   iter = ITER,
   warmup = 1000,
   chains = CHAINS,
   cores = CORES,
   control = CONTROLS
)

# prior predictive check
pp_check(m3.1, nsamples=100)

## Add LOO for model comparison
m3.1 <- add_criterion(m3.1, criterion = c("waic","loo"))



### MODEL 2

# prior model
m3.2_prior <- brm(
  formula3.2,
  data = subset(d3, ID %in% sub),
  family = gaussian,
  prior = prior,
  sample_prior = "only",
  chains = CHAINS,
  iter = ITER,
  control = CONTROLS
)

# prior predictive check
pp_check(m3.2_prior, nsamples=100)


# posterior model
m3.2 <- brm(
  formula3.2,
  data = d3, # subset(d3, ID %in% sub)
  family = gaussian,
  prior = prior,
  sample_prior = TRUE,
  chains = CHAINS,
  warmup = 1000,
  iter = ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.2, nsamples=100)

## Add LOO for model comparison
m3.2 <- add_criterion(m3.2, criterion = c("waic","loo"))

#save(m3.1, m3.2, file = "Models/Exp3_ErrorModels")


```

Report models (m3.1 is the best of them)
```{r}
# Summaries
summary(m3.1)
summary(m3.2)

# Model comparison
compare1 <- loo_compare(m3.1, m3.2)
weights1 <- loo_model_weights(m3.1, m3.2) 

compare1
weights1




# Summarizing and visualizing results
hypothesis(m3.1,"generationL<0")
hypothesis(m3.2,"generationL<0")
hypothesis(m3.2,"IgenerationL<0")

# Estimates with more decimals
fixef(m3.1)
fixef(m3.2)

# Plot models
plot(hypothesis(m3.1,"generationL<0"))
plot(hypothesis(m3.2,"generationL<0"))
conditional_effects(m3.1)
conditional_effects(m3.2)

plot(conditional_effects(m3.1), points=T, rug=T)[[1]] + theme_classic()

# Plot trace plots
plot(m3.1)
plot(m3.2)

```


model 3.1 (best from Loo comparison)
model 3.2 

--> proceed with the most simple (bad) model by adding chain as fixed effect to control for it + Interaction model.



## Control 


### Control analysis for Chain
Note: controlling for site did not improve the model in Tylén et al.

Pre-process the data --> make new chain column
```{r}

# Make chain group column
d3$chainGroup <- NA
for (i in (1:nrow(d3))){
  
  if (d3$ID[i] <= 5) {
    d3$chainGroup[i] <- 1
    
  } else if (d3$ID[i] > 5 & d3$ID[i] < 11) {
    d3$chainGroup[i] <- 2
    
  } else if (d3$ID[i] > 10 & d3$ID[i] < 16) {
    d3$chainGroup[i] <- 3
    
  } else if (d3$ID[i] > 15 & d3$ID[i] < 21) {
    d3$chainGroup[i] <- 4
    
  } else if (d3$ID[i] > 20 & d3$ID[i] < 26) {
    d3$chainGroup[i] <- 5
    
  } else if (d3$ID[i] > 25 & d3$ID[i] < 31) {
    d3$chainGroup[i] <- 6
    
  } else if (d3$ID[i] > 30 & d3$ID[i] < 36) {
    d3$chainGroup[i] <- 7
    
  } else if (d3$ID[i] > 35 & d3$ID[i] < 41) {
    d3$chainGroup[i] <- 8
    
  } else if (d3$ID[i] > 40 & d3$ID[i] < 46) {
    d3$chainGroup[i] <- 9
    
  } else if (d3$ID[i] > 45 & d3$ID[i] < 51) {
    d3$chainGroup[i] <- 10
    
  } else if (d3$ID[i] > 50 & d3$ID[i] < 56) {
    d3$chainGroup[i] <- 11
    
  } else { # In case we add 56-61 (which had the pictures from group 1)
    d3$chainGroup[i] <- 1
  }
}


# move df to old df
d3$chain <- as.character(d3$chain)
d3_old <- d3

# Create empty df
d3 <- d3[0,]

# Use chain group column to loop through them one by one and overwrite chains with a dummy.
for (i in 1:length(unique(d3_old$chainGroup))){
  
  # subset data
  sub <- d3_old %>% subset(chainGroup == i)
  
  sub$chainD <- sub$chain
  
  # change chains to a dummy by chaingroup
  sub$chainD <- as.numeric(as.factor(sub$chainD))
  sub$chainD <- sub$chainD - 1
  
  # Combine with premade empty dataframe
  if (nrow(d3) == 0) {
    d3 <- sub
    } else {
        d3 <- rbind(d3, sub)}

}




# change chain to be a factor
d3$chainD <- as.factor(d3$chainD)

```

Run 2 models controlling for Chain (fixed affect and interaction effect models)
```{r}

# define models including chain (corresponding to site - Blombos & Diepkloof)
formula3.3 <-  bf(MSE_s ~ 1 + generationL + chainD + 
                    (1 + generationL + chainD | ID) + 
                    (1 | drawing))

formula3.4 <-  bf(MSE_s ~ 1 + generationL * chainD + 
                    (1 + generationL * chainD | ID) + 
                    (1 | drawing))


### Model 5
m3.3 <- brm(
  formula3.3,
  data = d3,
  family = gaussian,
  prior = prior,  
  sample_prior = TRUE,
  chains = CHAINS,
  warmup = 1000,
  cores = CORES,
  iter=ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.3, nsamples=100)

# Add LOO for model comparison
m3.3 <- add_criterion(m3.3, criterion = c("waic","loo"))



### Model 5
m3.4 <- brm(
  formula3.4,
  data = d3,
  family = gaussian,
  prior = prior,  
  sample_prior = TRUE,
  chains = CHAINS,
  warmup = 1000,
  cores = CORES,
  iter=ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.4, nsamples=100)

## Add LOO for model comparison
m3.4 <- add_criterion(m3.4, criterion = c("waic","loo"))

```

Report the winning model
```{r}
# Summary
summary(m3.3)
summary(m3.4)


# Model comparison
compare2 <- loo_compare(m3.1, m3.2, m3.3, m3.4)
weights2 <- loo_model_weights(m3.1, m3.2, m3.3, m3.4)

compare2
weights2


# Summarizing and visualizing results
hypothesis(m3.3,"generationL<0")
hypothesis(m3.4,"generationL<0")

hypothesis(m3.4,"chainD1 < 0")
hypothesis(m3.4,"generationL:chainD1 > 0")


# Estimates with more decimals
fixef(m3.3)
fixef(m3.4)

# Plot models
plot(hypothesis(m3.3,"generationL<0"))
plot(hypothesis(m3.4,"generationL<0"))
conditional_effects(m3.3)
conditional_effects(m3.4)


# Plot trace plots
plot(m3.3)
plot(m3.4)

```

When controlling for chain by adding it as a fixed effect interacting with generation, we found a small negative effect of generation in the hypothesised direction.
Furthermore, we found a significant effect of the interaction between generation and chain --> when accounting for the fact that the symbols originated from two different chain groups (different symbols reproduced by different people), we unmasked a significant effect of generation in the hypothesised direction.


### Compare when excluding problematic pictures (to be done)

list of problematic stimuli (subjectively selected by Aske)
```{r}
problemPics <- c("i3_c104_g8_id16", "i3_c104_g4_id12", "i3_c104_g1_id9",
                 "i4_c102_g4_id15", "i4_c102_g8_id315", "i4_c102_g1_",
                 "i4_c108_g4_id48", "i4_c108_g8_id52", "i4_c108_g1_id45",
                 "i3_c103_g8_id8", "i3_c103_g4_id4", "i3_c103_g1_id1")

d3sub <- subset(d3, !drawing %in% problemPics)

```

Run on model 3.1 and 3.2 --> I call them 3.5 and 3.6
```{r}


### MODEL 5 (without problematic pictures)

# posterior model
m3.5 <- brm(
  formula3.1,
  data = subset(d3, !drawing %in% problemPics), 
  family = gaussian,
  prior = prior,
  sample_prior = TRUE,
  warmup = 1000,
  chains = CHAINS,
  iter = ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.5, nsamples=100)

## Add LOO for model comparison
m3.5 <- add_criterion(m3.5, criterion = c("waic","loo"))




### MODEL 6 (without problematic pictures)

# posterior model
m3.6 <- brm(
  formula3.2,
  data = subset(d3, !drawing %in% problemPics), 
  family = gaussian,
  prior = prior,
  sample_prior = TRUE,
  chains = CHAINS,
  warmup = 1000,
  iter = ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.6, nsamples=100)

## Add LOO for model comparison
m3.6 <- add_criterion(m3.6, criterion = c("waic","loo"))



# Compare to the original models
compare3 <- loo_compare(m3.5, m3.6)
weights3 <- loo_model_weights(m3.5, m3.6) 
weights3

# Summarizing and visualizing results
summary(m3.5)
summary(m3.6)


```

Run on model 3.3 and 3.4 --> I call them 3.7 and 3.8
```{r}

### MODEL 7 (without problematic pictures)

# posterior model
m3.7 <- brm(
  formula3.3,
  data = subset(d3, !drawing %in% problemPics), 
  family = gaussian,
  prior = prior,
  sample_prior = TRUE,
  warmup = 1000,
  chains = CHAINS,
  iter = ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.7, nsamples=100)

## Add LOO for model comparison
m3.7 <- add_criterion(m3.7, criterion = c("waic","loo"))




### MODEL 8 (without problematic pictures)

# posterior model
m3.8 <- brm(
  formula3.4,
  data = subset(d3, !drawing %in% problemPics), 
  family = gaussian,
  prior = prior,
  sample_prior = TRUE,
  chains = CHAINS,
  warmup = 1000,
  iter = ITER,
  control = CONTROLS
)

# posterior predictive check
pp_check(m3.4_prior, nsamples=100)

## Add LOO for model comparison
m3.8 <- add_criterion(m3.8, criterion = c("waic","loo"))





# Compare to the original models
exp3_compare4 <- loo_compare(m3.5, m3.7, m3.8)
exp3_weights4 <- loo_model_weights(m3.5, m3.7, m3.8) 

exp3_compare4
exp3_weights4


# Summarizing and visualizing results'
summary(m3.3)
summary(m3.4)

q <- 3

summary(m3.7)
summary(m3.8)
fixef(m3.8)


summary(m3.4)
hypothesis(m3.8,"generationL<0")
hypothesis(m3.8,"chainD1<0")
hypothesis(m3.8,"generationL:chainD1>0")


```


## THE FINAL MODEL TO BE REPORTED
```{r}

# summary
summary(m3.4)

# Model comparison LOOIC
compare1 # m3.1 (simple model) is better than quadratic model
weights1

#model comparison of the best simple model and models controling for Chain
compare3.5 <- loo_compare(m3.1, m3.3, m3.4)
weights3.5 <- loo_model_weights(m3.1, m3.3, m3.4)

compare5.5
weights5.5


# Hypothesis testing
hypothesis(m3.4,"generationL<0")
hypothesis(m3.4,"chainD1 < 0")
hypothesis(m3.4,"generationL:chainD1 > 0")


# Estimates with more decimals
fixef(m3.4)


# Plot model
plot(hypothesis(m3.8,"generationL<0"))
conditional_effects(m3.4)


# Plot trace plot
plot(m3.8)


```



# Plots
```{r}

library(pacman)
p_load(extrafont)

font_import(pattern="[T/t]imes")
loadfonts(device="win")


## Plot for the main manuscript

dS_agg <- d3 %>% group_by(ID, generationL) %>% 
  summarise(
    meanError = mean(MSE_s)
)

dS_agg$ID = as.factor(dS_agg$ID)


Exp3_MainPlot <- ggplot(dS_agg, aes(generationL, meanError)) + 
  geom_line(aes(group=ID,color=ID),alpha=0.6) +
  geom_point(aes(group=ID,color=ID),alpha=0.6)+
  geom_smooth(method=lm, color = "red") +
  scale_color_discrete(guide=FALSE) +
  scale_x_continuous(breaks = c(-1,0,1),labels=c("1st","4th","8th")) +
  
  # THEMES
  theme_grey() +
  theme(
    text = element_text(family = "Times New Roman"),
    legend.position="top",
    plot.subtitle=element_text(face="italic",size=14,colour="grey40"),
    plot.title=element_text(size=21,face="bold")) +
  labs(title="\nExperiment 3",
       subtitle="Memorability",
       x=expression("Generation"),
       y=expression("Average stimulus reproduction error")) +
  NULL

Exp3_MainPlot

#ggsave("Exp3_MainPlot.svg", plot = Exp3_MainPlot)

```















