---
title: "experiment1"
author: "Aske & Rebecca"
date: "6 October 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

EXPERIMENT 1 - SALIENCY

# Prepare data

Load packages, data, wd etc. 
```{r}
# packages
pacman::p_load(tidyverse,lmerTest,ggbeeswarm, lme4, dplyr, brms)

citation("brms")
citation("tidyverse")
citation("dplyr")



# WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp1_data")

```

Collect data from several csv files
```{r}

# import data from several csv files
fileList1 <- list.files(full.names = T, pattern =".csv")
d1 <- map_dfr(fileList1,read_csv)

```

Include ID 56-61
```{r}
# setwd
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/Collected data/exp1_data/1-6")

# import data from several csv files
fileList1.1 <- list.files(full.names = T, pattern =".csv")
d1.1 <- map_dfr(fileList1.1,read_csv)

# Change ID before merging
d1.1$Participant <- d1.1$Participant + 55

d1 <- rbind(d1, d1.1) %>% 
  rename(ID = Participant, RT = Reaction_time)
  

# Create csv in shared folder
write.csv(d1,"C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske/exp1_collected_data.csv", row.names = FALSE)

```

Data tidying
```{r}

# Set new WD
setwd("C:/Users/askes/Dropbox/Bachelor - Aske og Rebecca/data_analysis/Bachelor_Rebecca_Aske")
d1 <- read_csv("exp1_collected_data.csv")

# Change column categories
d1 <- d1 %>% 
  mutate(
    ID = as.factor(ID),
    Age = as.factor(Age),
    Gender = as.factor(Gender),
    Block = as.factor(Block),
    Drawing = as.factor(Drawing),
    Chain = as.factor(Chain),
    Generation = as.factor(Generation),
    Response = as.factor(Response),
    Correct_response = as.factor(Correct_response),
    Correctness = as.factor(Correctness)
  )


# Check datapoints per IDs
nrow(d1) - (144 * length(unique(d1$ID)))
check <- d1 %>% group_by(ID, Block) %>% summarize(n()) 
names(check)[3] <- "count"
check %>% filter(count > 48) 
# We have 4 too many rows


# Remove NaN's, negative RT's, and wrong answers
d1 <- d1 %>% subset(RT > 0) %>% subset(Correctness != "NaN") %>% subset(Correctness == 1)


# Change values of generation in new colum to be -1, 0 or 1.
d1$GenerationNumber <- as.numeric(gsub("g", "", d1$Generation))
d1$GenerationNumber <- ifelse(d1$GenerationNumber == 1, -1, d1$GenerationNumber)
d1$GenerationNumber <- ifelse(d1$GenerationNumber == 4, 0, d1$GenerationNumber)
d1$GenerationNumber <- ifelse(d1$GenerationNumber == 8, 1, d1$GenerationNumber)


# Make a standardized RT column
d1$RT_standard <- d1$RT/max(d1$RT)
hist(d1$RT_standard)
hist(d1$RT)

```

Data visualisation
```{r}

#Plot of the data
ggplot(d1, aes(Generation, RT)) + 
  geom_point() +
  geom_smooth(method=lm,aes(group = ID, color=ID), se=F) +
  theme_classic()


# Double-checking the RTs
ggplot(d1,aes(RT)) + 
  geom_density(alpha=0.3)

# RTs standardized
ggplot(d1,aes(RT_standard)) + 
  geom_density(alpha=0.3)


## By ID
ggplot(d1,aes(RT)) + 
  geom_density(aes(fill=ID,group=ID,color=ID),alpha=0.3) +
  guides(fill=FALSE,group=FALSE,color=FALSE) +
  NULL


ggplot(subset(d1,Correctness == 1), aes(GenerationNumber,RT)) +
  geom_violin() +
  geom_beeswarm(aes(color=Generation))

```



# Run Models (Bayesian)

## Linear and quadratic regression models (lognormal) RT ~ generation (on all data)
```{r}
# define chains, iter, and controls
CHAINS = 4
CORES = 4
ITER = 4000
WARMUP = 1000

CONTROLS = list(
  max_treedepth = 20,
  adapt_delta=0.99)

# Choose ID's for subsettign the data
sub <- c(1, 2, 3, 4, 5, 6)

# Models (linear)
formula1.1 <- bf(RT ~ 1 + GenerationNumber + (1 + GenerationNumber | ID) + (1|Drawing))

# Seems lognormally distributed constrained between 0 and 15.
hist(d1$RT, breaks = 70, density = 20)

get_prior(formula1.1, d1, family = lognormal())

# set priors
prior1.1 <- c(
  prior(normal(0.5, 0.2), class = Intercept),
  prior(normal(0, 0.5), class = b),
  prior(normal(0, 0.2), class = sd),
  prior(normal(0, 0.2), class = sigma),
  prior(lkj(5), class = cor))


### MODEL 1.1

# prior
m1.1_prior <- brm(
   formula1.1,
   subset(d1, ID %in% sub),
   family = lognormal(),
   prior = prior1.1,
   sample_prior = "only",
   chains = 1,
   cores = 1,
   iter = 200,
   warmup = 100
)

pp_check(m1.1_prior, nsamples=100)

# posterior
m1.1 <- brm(
  formula1.1,
  d1,
  family = lognormal(),
  prior = prior1.1,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER,
  warmup = WARMUP
)

pp_check(m1.1, nsamples=100)


### MODEL 2

# formula
formula1.2 <- bf(RT ~ 1 + GenerationNumber + I(GenerationNumber)^2 + 
                   (1 + GenerationNumber  + I(GenerationNumber)^2 | ID) + (1|Drawing))

# prior
m1.2_prior <- brm(
   formula1.2,
   subset(d1, ID %in% sub),
   family = lognormal(),
   prior = prior1.1,
   sample_prior = "only",
   chains = 1,
   cores = 1,
   iter = 200,
   warmup = 100
)

pp_check(m1.2_prior, nsamples=100)

# posterior
m1.2 <- brm(
  formula1.2,
  d1,
  family = lognormal(),
  prior = prior1.1,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER,
  warmup = WARMUP
)


pp_check(m1.2, nsamples=100)

```

Model 1.1. and 1.2 comparison
```{r}
m1.1 <- add_criterion(m1.1, criterion = c("bayes_R2", "loo"))
m1.2 <- add_criterion(m1.2, criterion = c("bayes_R2", "loo"))

exp1_compare1 <- loo_compare(m1.1, m1.2)
exp1_weights1 <- loo_model_weights(m1.1,m1.2)


exp1_compare1
exp1_weights1

```

Report best model
```{r}

# summaries
summary(m1.1)
summary(m1.2)

# hypothesis testing
hypothesis(m1.1, "GenerationNumber < 0") # no significant effect, however effect in the right direction
plot(hypothesis(m1.1, "GenerationNumber < 0"))




# remember trace plots etc.
plot(m1.1)
plot(m1.2)

```



## Control for Chain (model 1.3 and 1.4)

Pre-process the data --> make new chain column
```{r}
# Make chain group column
d1$chainGroup <- NA
d1$ID <- as.numeric(d1$ID)
for (i in (1:nrow(d1))){
  
  if (d1$ID[i] <= 5) {
    d1$chainGroup[i] <- 1
    
  } else if (d1$ID[i] > 5 & d1$ID[i] < 11) {
    d1$chainGroup[i] <- 2
    
  } else if (d1$ID[i] > 10 & d1$ID[i] < 16) {
    d1$chainGroup[i] <- 3
    
  } else if (d1$ID[i] > 15 & d1$ID[i] < 21) {
    d1$chainGroup[i] <- 4
    
  } else if (d1$ID[i] > 20 & d1$ID[i] < 26) {
    d1$chainGroup[i] <- 5
    
  } else if (d1$ID[i] > 25 & d1$ID[i] < 31) {
    d1$chainGroup[i] <- 6
    
  } else if (d1$ID[i] > 30 & d1$ID[i] < 36) {
    d1$chainGroup[i] <- 7
    
  } else if (d1$ID[i] > 35 & d1$ID[i] < 41) {
    d1$chainGroup[i] <- 8
    
  } else if (d1$ID[i] > 40 & d1$ID[i] < 46) {
    d1$chainGroup[i] <- 9
    
  } else if (d1$ID[i] > 45 & d1$ID[i] < 51) {
    d1$chainGroup[i] <- 10
    
  } else if (d1$ID[i] > 50 & d1$ID[i] < 56) {
    d1$chainGroup[i] <- 11
    
  } else { # In case we add 56-61 (which had the pictures from group 1)
    d1$chainGroup[i] <- 1
  }
}


# move df to old df
d1$Chain <- as.character(d1$Chain)
d1_old <- d1

# Create empty df
d1 <- d1[0,]

# Use chain group column to loop through them one by one and overwrite chains with a dummy.
for (i in 1:length(unique(d1_old$chainGroup))){
  
  # subset data
  sub <- d1_old %>% subset(chainGroup == i)
  
  sub$chainD <- sub$Chain
  
  # change chains to a dummy by chaingroup
  sub$chainD <- as.numeric(as.factor(sub$chainD))
  sub$chainD <- sub$chainD - 1
  
  # Combine with premade empty dataframe
  if (nrow(d1) == 0) {
    d1 <- sub
    } else {
        d1 <- rbind(d1, sub)}

}


# change chain to be a factor
d1$chainD <- as.factor(d1$chainD)

```

Run the models
```{r}

# Models (linear)
formula1.3 <-  bf(RT ~ 1 + GenerationNumber + chainD + (1 + GenerationNumber + chainD | ID) + (1 | Drawing))

get_prior(formula1.3, d1, family = lognormal())

# set priors
prior1.3 <- c(
  prior(normal(0.5, 0.2), class = Intercept),
  prior(normal(0, 0.5), class = b),
  prior(normal(0, 0.2), class = sd),
  prior(normal(0, 0.2), class = sigma),
  prior(lkj(5), class = cor))

# Choose ID's for subsettign the data
sub <- c(1, 2, 3, 4, 5, 6)

### MODEL 1.1

# prior
m1.3_prior <- brm(
   formula1.3,
   data = subset(d1, ID %in% sub),
   family = lognormal(),
   prior = prior1.3,
   sample_prior = "only",
   chains = 1,
   cores = 1,
   iter = 200,
   warmup = 100
)

pp_check(m1.3_prior, nsamples=100)

# posterior
m1.3 <- brm(
  formula1.3,
  d1,
  family = lognormal(),
  prior = prior1.3,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER,
  warmup = WARMUP
)

pp_check(m1.3, nsamples=100)


### MODEL 4

# formula
formula1.4 <-  bf(RT ~ 1 + GenerationNumber*chainD + (1 + GenerationNumber*chainD | ID) + (1 | Drawing))

# prior
m1.4_prior <- brm(
   formula1.4,
   subset(d1, ID %in% sub),
   family = lognormal(),
   prior = prior1.3,
   sample_prior = "only",
   chains = 1,
   cores = 1,
   iter = 200,
   warmup = 100
)

pp_check(m1.4_prior, nsamples=100)

# posterior
m1.4 <- brm(
  formula1.4,
  d1,
  family = lognormal(),
  prior = prior1.3,
  sample_prior = T,
  chains = CHAINS,
  cores = CORES,
  iter = ITER,
  warmup = WARMUP
)


pp_check(m1.4, nsamples=100)

```

Model 1.3. and 1.4 comparison
```{r}
m1.3 <- add_criterion(m1.3, criterion = c("bayes_R2", "loo"))
m1.4 <- add_criterion(m1.4, criterion = c("bayes_R2", "loo"))

exp1_compare2 <- loo_compare(m1.3, m1.4)
exp1_weights2 <- loo_model_weights(m1.3, m1.4)

exp1_compare3 <- loo_compare(m1.1, m1.3, m1.4)
exp1_weights3 <- loo_model_weights(m1.1, m1.3, m1.4)


exp1_weights3

```

Report best model
```{r}

# summaries
summary(m1.3)
summary(m1.4)

# hypothesis testing
hypothesis(m1.3, "GenerationNumber < 0") 
plot(hypothesis(m1.3, "GenerationNumber < 0"))

hypothesis(m1.4, "GenerationNumber < 0") 
plot(hypothesis(m1.4, "GenerationNumber < 0"))


# remember trace plots etc.
plot(m1.3)
plot(m1.4)

```


## Control for problematic pictures and ID
```{r}

# control for problematic pictures. list of problematic stimuli (subjectively selected by Aske)
problemPics <- c("i3_c104_g8_id16", "i3_c104_g4_id12", "i3_c104_g1_id9",
                 "i4_c102_g4_id15", "i4_c102_g8_id315", "i4_c102_g1_",
                 "i4_c108_g4_id48", "i4_c108_g8_id52", "i4_c108_g1_id45",
                 "i3_c103_g8_id8", "i3_c103_g4_id4", "i3_c103_g1_id1")

# update the 4 first models (those we will look at)
m1.5 <- update(m1.1, newdata = subset(d1, !Drawing %in% problemPics))
m1.6 <- update(m1.2, newdata = subset(d1, !Drawing %in% problemPics))
m1.7 <- update(m1.3, newdata = subset(d1, !Drawing %in% problemPics))
m1.8 <- update(m1.4, newdata = subset(d1, !Drawing %in% problemPics))

# Do the same excludign problemtic ID's as well

# ID: 10, 11, 12, 13, 30, 42, 43, 56(1) exclude block 1.
probID <- c(10, 11, 12, 13, 30, 42, 43, 56)
d1_probID <- d1 %>% subset(!ID %in% probID) %>% subset(!Drawing %in% problemPics)

ID56 <- d1 %>% subset(!Drawing %in% problemPics) %>% filter(ID == 56 & Block != 1)
d1_probID <- rbind(d1_probID, ID56)

# update models
m1.9 <- update(m1.1, newdata = d1_probID)
m1.10 <- update(m1.2, newdata = d1_probID)
m1.11 <- update(m1.3, newdata = d1_probID)
m1.12 <- update(m1.4, newdata = d1_probID)


```

Model comparison
```{r}

m1.5 <- add_criterion(m1.5, criterion = c("bayes_R2", "loo"))
m1.6 <- add_criterion(m1.6, criterion = c("bayes_R2", "loo"))
m1.7 <- add_criterion(m1.7, criterion = c("bayes_R2", "loo"))
m1.8 <- add_criterion(m1.8, criterion = c("bayes_R2", "loo"))
m1.9 <- add_criterion(m1.9, criterion = c("bayes_R2", "loo"))
m1.10 <- add_criterion(m1.10, criterion = c("bayes_R2", "loo"))
m1.11 <- add_criterion(m1.11, criterion = c("bayes_R2", "loo"))
m1.12 <- add_criterion(m1.12, criterion = c("bayes_R2", "loo"))

# compare models excluding prob pics
exp1_compare4 <- loo_compare(m1.5, m1.6)
exp1_weights4 <- loo_model_weights(m1.5, m1.6)

exp1_compare5 <- loo_compare(m1.7, m1.8)
exp1_weights5 <- loo_model_weights(m1.7, m1.8)

exp1_compare6 <- loo_compare(m1.5, m1.7, m1.8)
exp1_weights6 <- loo_model_weights(m1.5, m1.7, m1.8)

exp1_compare4
exp1_weights4


# Compare models ecluding both problematic ID's and pics
exp1_compare7 <- loo_compare(m1.9, m1.10)
exp1_weights7 <- loo_model_weights(m1.9, m1.10)

exp1_compare8 <- loo_compare(m1.11, m1.12)
exp1_weights8 <- loo_model_weights(m1.11, m1.12)

exp1_compare9 <- loo_compare(m1.9, m1.11, m1.12)
exp1_weights9 <- loo_model_weights(m1.9, m1.11, m1.12)


```



# Report overall best model (wednesday) --> without prob pics.
```{r}
# Summaries

# most simple model (the one to be reported) 
summary(m1.5)
plot(m1.5)
pp_check(m1.5_prior, nsamples=100)

fixef(m1.5)

# recalculate estimates to seconds to be more interpretable
exp(0.34)

exp(0.34-0.01)

exp(0.34-0.01-0.01)

exp1_compare6
exp1_weights6

hypothesis(m1.5, "GenerationNumber < 0")
plot(hypothesis(m1.5, "GenerationNumber < 0"))


# - linear regression model contrlling for ChainGroup
summary(m1.7)
hypothesis(m1.7, "GenerationNumber < 0")
hypothesis(m1.7, "chainD1 < 0")



# with interaction
summary(m1.8)


# without prob ID's
summary(m1.9)
summary(m1.10)
summary(m1.11)
summary(m1.12)




```




# Plots
```{r}

# Create summary dataset for visualisation

plotSum1 <- d1 %>% group_by(ID, GenerationNumber) %>% summarise(
  RTmean = mean(RT)
)

plotSum1$ID <- as.factor(plotSum1$ID)
plotSum1$GenerationNumber <- as.numeric(plotSum1$GenerationNumber)

Exp1_MainPlot <- ggplot(plotSum1, aes(GenerationNumber, RTmean)) + 
  geom_line(aes(group=ID,color=ID),alpha=0.6) +
  geom_point(aes(group=ID,color=ID),alpha=0.6)+
  geom_smooth(method=lm, color = "red") +
  scale_color_discrete(guide=FALSE) +
  scale_x_continuous(breaks = c(-1,0,1), labels=c("1st","4th","8th")) +

  # THEMES
  theme_grey() +
  theme(
    text = element_text(family = "Times New Roman"),
    legend.position="top",
    plot.subtitle=element_text(face="italic",size=14,colour="grey40"),
    plot.title=element_text(size=21,face="bold")) +
  labs(title="\nExperiment 1",
       subtitle="Saliency",
       x=expression("Generation"),
       y=expression("Time to Emerge (seconds)")) +
  NULL
  
Exp1_MainPlot


```
















